Thesis Draft
============

Hierarchy of information in data analysis paper 
-----------------------------------------------
- Title / Author list
- Abstract
- Body / Results
- Supplementary Materials (the gory details)
- Code/Data (really gory details)

Steps in a data analysis
------------------------
### Deﬁne the question

- Do past Kickstarter projects' successes/failures influence future successes/ failures Kickstarter projects?

- (more generic) predicting of Kickstarter project success

### Deﬁne the ideal data set

Ideally, we would have information about all video game projects ever crowd-funded, including economical and sociological characteristics, such as: the total amount raised, the number of backers, the project timeframe, all interconnections between backers (are they know each other? have they influenced each other decision about project contribution?), also every news/tweet ever made concerning video game projects or project owners, or backers themselves that could be used in decision making related to backing up the project. 

### Determine what data you can access

- from Kickstarter: a main page about project, including comments, and videos, the total number of backers and collected funds, optionally links to backers social (facebook page or/and tweeter stream, or/and personal web page), 

- from Kicktraq: the number of backers and their contributions per day wihtin a project lifetime, 

- (optionally) from Facebook backer page: messages about backing the project if configured as publically available,

- (optionally) from Facebook project page: all the information about people how support the project ('liked' it), messages, comments, etc.

- (optionally) from Youtube project videos: comments, total number of views

### Obtain the data
We can obtain the data from websites, like: Kickstarter, Kicktraq, Facebook and Youtube. Due to the fact that all of them have different API, the procedure of collecting the raw data is different in each case. The initial set is to retrieve all relevant projects

#### Data collection from Kickstarter
Kickstarter data is hierarchical by nature and API follows the same concept. The primary entity is ***a project*** owned by ***the owner*** and supported by ***backers*** with ***funds***. The primary goal is to collect enough funds to launch a product or a service on the market which is refered to as ***a successful project***. The incentive of a backer is usually a faster access to a new produt/service for a cheaper price in comparision to a launch price, the risk is that the release is not going to take place because the collected amount of funds is insufficient.

All ***Video Game*** finished projects are in scope: successful, unsuccessful and cancelled. We shall navigate to a corresponding result page using Kickstarter. Due to pagination, we have to download page by page. The last page will have no results.
```{r, }
require(jsonlite)

# retrieve and merge until gives no more results
rawprojectsdir <- "./raw/kickstarter/projects/"
if (!file.exists(rawprojectsdir)) {
    dir.create(rawprojectsdir,recursive=TRUE,showWarnings=FALSE)
    doBreak <- FALSE
    i <- 1
    while(!doBreak) {
        download.file(
            paste0("https://www.kickstarter.com/discover/advanced?page=",i,"&category_id=35&sort=launch_date"), 
            paste0("./raw/kickstarter/projects/page",i,".json"),
            method = "curl", extra = c("--header","accept:application/json"))
        # parse to JSON
        thisPage <- fromJSON(paste0("./raw/kickstarter/projects/page",i,".json"))
        doBreak <- length(thisPage$projects) == 0
        inc(i) <- 1
    }
}
```

Having all relevant projects downloaded, we may proceed with the next step, which is the retrieval of main pages of each project. We will create a subdirectory for each project and store main page file there. We will fetch raw html format and extract a project representation in JSON format from an embedded script element.  

```{r}
require(jsonlite)
require(XML)
require(stringr)

downloadKickstarterProjectPage <- function(pid,url,rootdir) {
    projectdir <- paste0(rawprojectsdir,'/',pid,'/')
    if (!file.exists(projectdir)) {
        message(paste("Processing Kickstarter project",pid,"(",url,")."))
        dir.create(projectdir, recursive=TRUE, showWarnings=FALSE)
        htmlfile <- paste0(projectdir,pid,'.html')
        jsonfile <- paste0(projectdir,pid,'.json')
        download.file(url,htmlfile,method="curl")
        html <- htmlTreeParse(htmlfile,useInternalNodes=T)
        system(paste("gzip",htmlfile))
        script <- xpathSApply(html, "//script[contains(./text(),'window.current_project')]" ,xmlValue)
        matcher <- str_match(script,"window.current_project.*\"(\\{.*\\})\"")
        if (!is.na(matcher[,2])) {
            json <- gsub('&quot;','"',matcher[,2])
            json <- gsub('\\\\\"','\\"',json)
            cat(prettify(toJSON(fromJSON(json))),file = jsonfile)
            system(paste("gzip",jsonfile))
        }
    }
    else {
        message(paste("Skipping Kickstarter project",pid,"(",url,"), because raw files already exist."))
        unlink(projectdir, recursive=T)
    }
}

rawprojectsdir <- "./raw/kickstarter/projects"
pagefiles <- list.files(rawprojectsdir,pattern="page.*.json", full.names=T)
for(pagefile in pagefiles) {
    message(paste("Processing page file",pagefile))
    json <- fromJSON(pagefile)
    pid <- json$projects$id
    url <- json$projects$urls$web$project
    mapply(downloadKickstarterProjectPage,
               pid,url,MoreArgs = list(rootdir = rawprojectsdir))
}
```

#### Data collection from Indiegogo
Indiegogo does provide one REST service which allows to retrieve a project JSON file knowing the project id. However, there is no service that gives project ids per category, and we have to loop over all indiegogo projects and retrieve all of them.

Another problem is that the project category is being interpreted more freely than on Kickstarter - gaming category includes computer games, board games, even requests for support to buy new hardware. A classification based on project description is required to filter out projects that are not related to computer games.

```{r}

downloadIndiegogoProjectPage <- function(pid,rootdir) {
    projectdir <- paste0(rootdir,'/',pid)
    url <- paste0("https://www.indiegogo.com/projects",'/',pid,'/')
    if (!file.exists(projectdir)) {
        message(paste("Processing Indiegogo project",pid,"(",url,")."))
        dir.create(projectdir, recursive=TRUE, showWarnings=FALSE)
        jsonfile <- paste0(projectdir,'/',pid,".json")
        download.file(url, jsonfile,
            method = "curl", extra = c("--header","accept:application/json"))
        if (file.info(jsonfile)$size > 5) {
        system(paste("gzip",jsonfile))
        htmlfile <- paste0(projectdir,'/',pid,".html")
        download.file(paste0(url,"widget"),htmlfile,
            method = "curl", extra = c("--header","accept:text/html"))
        system(paste("gzip",htmlfile))
        }
        else {
            message(paste("Skipping Indiegogo project",pid," because of empty JSON file. Probably project is a draft that never got published."))
            unlink(projectdir, recursive=T)
        }
    }
    else {
        message(paste("Skipping Indiegogo project",pid,"(",url,"), because raw files already exist."))        
    }
}

rawdir_indiegogo <- "./raw/indiegogo/projects/20001-30000"
pids <- c(20001:30000)
mapply(downloadIndiegogoProjectPage,
       pids,MoreArgs = list(rootdir = rawdir_indiegogo))
```

### Clean the data
In this chapter we are going to process every project JSON file and build one data table file with all the variables we can extract from that file.

#### Kickstarter data cleaning
The following information about the project can be easily extracted from a JSON project file:
- a title
- an owner
- a blurb (a short description)
- an owner location - a city and a country
- if the project was succesful
- the start and the finish/cancel date
- the duration of the project
- the total amount and currency of pledged
- the total number of backers
- the above target percentage
- if a project page contains a video link
- the list of rewards with the following metadata: a description, a value, a number of backers, an estimated delivery date
- the total number of provided comments
- the total number of project page updates

The exact names of currently used variables:
- pid (the unique project ID of Kickstarter)
- blurb (a short description displayed on project list)
- goal (the amount of money to collect for the project to be successful)
- pledged (he total amount of money pledged)
- state (a project current status [canceled / failed / live / successful / suspended])
- country (a country in which project is being developed)
- currency (a currency in which goal is set and pledges are collected)
- deadline (an end date of pledging)
- created_at (a project creation date)
- launched_at ( a date at which project page was made public and pledging was started) 
- backers_count (the total amount of backers)
- creator_id (Kickstarter unique user ID of the creator)
- creator_name (the name / nickname of the project creator)
- has_video (if the front page has at least one video embedded [TRUE / FALSE])
- location_id (Kickstarter unique city ID in which project is being developed)
- location_name (a city in which project is being developed)
- comments_count (the total amount of backer comments on project page [only backers can post comments])
- updates_count (the total amount of updates made to the project page by the creator)
- reward_value (the minimal amount of money needed to unlock given reward [backers may choose to pay more than value!])
- reward_desc (a reward description)
- reward_shipping_applicable (if the reward has shipping possibility)
- reward_backers_count (the total amount of backers who paid for that reward type)
- reward_backers_limit (how many rewards of that type is available [NA = no limit])

```{r}
require(jsonlite)
require(stringr)
require(data.table)

cleanAndProcessProject <- function(projectdir) {
    jsonfile <- paste0(projectdir,'/',basename(projectdir),".json")
    message(paste("Processing JSON project file",jsonfile))
    json <- fromJSON(jsonfile)
    reward_value <- json$rewards$minimum
    reward_backers_count <- json$rewards$backers_count
    if (is.null(reward_backers_count)) {
        reward_backers_count <- rep(NA,length(reward_value))
    }
    reward_shipping_applicable <- json$rewards$shipping_applicable
    if (is.null(reward_shipping_applicable)) {
        reward_shipping_applicable <- rep(NA,length(reward_value))
    }
    reward_backers_limit <- json$rewards$limit
    if (is.null(reward_backers_limit)) {
        reward_backers_limit <- rep(NA,length(reward_value))
    }
    data.table(pid=json$id, name=json$name, blurb=json$blurb, goal=json$goal, 
                      pledged=json$pledged, state=json$state, country=json$country, currency=json$currency,
                      deadline=structure(json$deadline,class=c('POSIXct')),
                      created_at=structure(json$created_at,class=c('POSIXct')),
                      launched_at=structure(json$launched_at,class=c('POSIXct')),
                      backers_count=json$backers_count, 
                      creator_id=json$creator$id, 
                      creator_name=json$creator$name,
                      has_video=!is.null(json$video$frame),
                      location_id=ifelse(is.null(json$location$id),NA,json$location$id), 
                      location_name=ifelse(is.null(json$location$displayable_name),
                                           NA,json$location$displayable_name),
                      comments_count=json$comments_count, 
                      updates_count=json$updates_count,
                      reward_value=reward_value,
                      reward_desc=json$rewards$reward,
                      reward_shipping_applicable=reward_shipping_applicable,
                      reward_backers_count=reward_backers_count,
                      reward_backers_limit=reward_backers_limit)
}

rawprojectsdir <- "./raw/kickstarter/projects"
cleandir <- "./clean/kickstarter"
projectsubdirs <- list.dirs(rawprojectsdir,full.names=T,recursive=F)
message(paste("Processing ",length(projectsubdirs)," JSON project files."))
dt <- data.table()
dt <- rbindlist(mapply(cleanAndProcessProject,projectsubdirs,SIMPLIFY=F))
write.csv(dt,file = paste0(cleandir,'/',"kickstarter_clean_withRewards.csv"))
udt <- unique(dt[,1:19,with=F],by=c(1))
write.csv(udt,file = paste0(cleandir,'/',"kickstarter_clean_noRewards.csv"))
```

### Exploratory data analysis

### Statistical prediction/modeling

### Interpret results

### Challenge results

### Synthesize/write up results

### Create reproducible code
